---
title: "Project 3 Variables Analysis"
author: "Wing. Modified by Iman Singh"
date: "10/31/2017"
output: html_document
---
## 1. Introduction

#### Exploration of variables 21-60 of the house train dataset to see if there are any that can be excluded.


```{r Initialization}
library(data.table)
library(ggplot2)
library(gridExtra)
library(tabplot)
library(lsr)
library(corrplot)
library(dplyr)
library(VIM)
library(caret)
library(RANN)
library(reshape2)
getwd()
house_train <- read.csv("~/Documents/NYC Data Science Academy/Project_3_Kaggle/train.csv", 
                        header = TRUE, 
                        na.strings = c("", "NA"),
                        stringsAsFactors = TRUE)
```

Create subset of house_train to include only variables 1-20 and 60-81:

```{r model_subset}
house_sub <- house_train[,c(21:61, 81)] 
```

## 2. Data Loading & Preparation


###Get levels of categorical features from data description:

```{r model factor_levels}
factorLevel <- list()
conn <- file("~/Documents/NYC Data Science Academy/Project_3_Kaggle/data_description.txt", open="r")
f <-readLines(conn)
for (line in f){
  if(!grepl("^[[:blank:]]", line) & grepl(": ", line)) {
    col_name <<- trimws(gsub(":.*", "", line))
  } else {
    level <- trimws(gsub("\t.*", "", line))
    if (level != "") {
      factorLevel[[col_name]] <- c(factorLevel[[col_name]], level)
    }
  }
}
close(conn)
unlink("~/Documents/NYC Data Science Academy/Project_3_Kaggle/Data_Descriptions.txt")

print(factorLevel)
print(names(factorLevel))
```

```{r check_levels}
## check if levels in description cover unique data values
for (varname in names(factorLevel)) {
  levelDiff <- setdiff(unique(house_train[[varname]]), 
                       factorLevel[[varname]])
  if(length(levelDiff)) {
    print(paste(varname, 
                paste(levelDiff, collapse = ", "), 
                sep = ": "))
  }
}
```

###Fix those levels that don't match with data
###Ignore "NA" as they will be considered as missing when converting categorical to factors.

```{r fix_levels}

unique(house_sub$MSZoning)
factorLevel$MSZoning
factorLevel$MSZoning[2] <- "C (all)"

unique(house_sub$Neighborhood)
factorLevel$Neighborhood
factorLevel$Neighborhood[13] <- "NAmes"

unique(house_sub$BldgType)
factorLevel$BldgType
factorLevel$BldgType[c(2,3,5)] <- c("2fmCon","Duplex","Twnhs")

unique(house_sub$Exterior2nd)
factorLevel$Exterior2nd
factorLevel$Exterior2nd[c(17,6,3)] <- c("Wd Shng","CmentBd","Brk Cmn")
```


###Get levels that only appear in the dataset
```{r get_levels}

for (varname in names(factorLevel)) {
  factorLevel[[varname]] <- intersect(factorLevel[[varname]],
                                      unique(house_sub[[varname]]))
}
```

###Convert column datatype to numeric / factor on training dataset:
```{r convert_types}
for (varname in names(house_sub)[-1]) {
  if (varname %in% names(factorLevel)) {
    house_sub[[varname]] <- factor(house_sub[[varname]], 
                                     levels = factorLevel[[varname]])
  } else {
    house_sub[[varname]] <- as.numeric(house_sub[[varname]])
  }
}
```

Save data frame:

```{r save_loaded}
house_sub$Id <- NULL
save(house_sub, file = "~/Documents/NYC Data Science Academy/Project_3_Kaggle/house_loaded.RData")
```

###create data frame of variable names and class type to copy to Excel (commented out code):
#### x  = data.frame(str(house_sub))
#### class_var = sapply(house_sub, class)
#### x2 = cbind(x, class_var)
####x2$colnames.house_sub. = NULL

## 3. Visualizations

### table plot all features on sortded SalePrice

```{r density_plots}
colMtx <- matrix(names(house_sub)[1:length(house_sub)], nrow = 6)
for (i in 1:ncol(colMtx)) {
  tableplot(house_sub, 
            select_string = c(colMtx[,i], "SalePrice"), 
            sortCol = "SalePrice", decreasing = TRUE, 
            nBins = 30)
}
```

### Correlations between Continuous Variables

```{r corrplot_numerical}
numeric_features <- names(house_sub)[sapply(house_sub, is.numeric)]
print(numeric_features)
#numeric_features <- numeric_features[-length(numeric_features)]
#print(numeric_features)
## correlation between continuous variables in training dataset - pearson
corr_numtran <- cor(house_sub %>% 
                      select(one_of(numeric_features, "SalePrice")), 
                    method = "pearson", 
                    use = "pairwise.complete.obs")

corrplot(corr_numtran, method = "color", order="hclust")

```

### Correlations between Ordinal Variables

```{r ordinal_features}
## ordinal features are those who contain one of the follow levels
ordinal_levels <- c("Flat", "ClyTile", "AsbShng", 
                    "BrkCmn", "Gd", "BrkTil", "GLQ",
                    "Floor")
ordinal_features <- character(0)
print(ordinal_features)
for(feature in names(house_sub)) {
  if(is.factor(house_sub[,feature]) && 
     length(intersect(ordinal_levels, levels(house_sub[,feature])))) {
    ordinal_features <- c(ordinal_features, feature)
  }
}

print(ordinal_features)

corr_ordtran <- cor(data.matrix(house_sub %>% 
                                  select(one_of(ordinal_features))), 
                    method = "kendall", 
                    use = "pairwise.complete.obs")
corrplot(corr_ordtran, method = "color", order="hclust")

```


```{r corrplot_nominal}
## Cramér's V is a measure of association between two nominal variables, giving a value between 0 and +1 (inclusive) NOT USED: ERROR IN RUNNING CHI-SQ APPROX.
cor.cramersV <- function(data) {
  cramersV(table(data[complete.cases(data),]))
}

nominal_features <- setdiff(names(house_sub), 
                            c(numeric_features, ordinal_features, "SalePrice"))
print(nominal_features)

```

### Correlations between Ordinal Variables vs. SalePrice

```{r corrplot_ordinal_2}
## coorelation between ordered categorical variables in training - spearman
cor.ordcnt <- function(data, x, y) {
  cor(as.numeric(data[[x]]), as.numeric(data[[y]]), 
                 method = "spearman", 
                 use = "pairwise.complete.obs")
}

corr_ordcnttran <- data.frame(Variable = ordinal_features,
                              Correlation = sapply(ordinal_features, 
                                                function(x) -cor.ordcnt(house_sub, x, "SalePrice")))

ggplot(corr_ordcnttran, aes(reorder(Variable, Correlation), Correlation)) + 
  geom_bar(stat = "identity") +
  coord_flip()
corr_ordcnttran

```

### Boxplots
```{r boxplots}
numeric_df <- house_sub %>% select(numeric_features)
#numeric_melt  <- melt(numeric_df)
ggplot(data = numeric_df, mapping = aes(x = value, y=)) + 
    geom_histogram(bins = 10) + facet_wrap(~variable, scales = 'free_x')

```

```{r boxplots}
ordinal_df <- house_sub %>% select(ordinal_features)
ordinal_melt  <- melt(ordinal_df)
ggplot(data = ordinal_melt, mapping = aes(x = value)) + 
    geom_boxplot() + facet_wrap(~variable, scales = 'free_x')

```

##4. Missingness. 

```{r missing_check}

## check missing values
col_missing <- names(house_sub)[colSums(is.na(house_sub)) > 0]
aggr(house_sub[,col_missing], prop = F, numbers = T)
Filter(function(x) x > 0, colSums(is.na(house_sub)))

miss_df <- colSums(house_sub[, c(1:40)] == '', na.rm = TRUE)
miss_df
```

``` {r na_sum}
sum(is.na(house_sub$YearRemodAdd))
sum(is.na(house_sub$RoofStyle))
sum(is.na(house_sub$RoofMatl))
sum(is.na(house_sub$Exterior1st))
sum(is.na(house_sub$Exterior2nd))
sum(is.na(house_sub$MasVnrType))
sum(is.na(house_sub$MasVnrArea))
sum(is.na(house_sub$ExterQual))
sum(is.na(house_sub$ExterCond))
sum(is.na(house_sub$Foundation))
sum(is.na(house_sub$BsmtQual))
sum(is.na(house_sub$BsmtCond))
sum(is.na(house_sub$BsmtExposure))
sum(is.na(house_sub$BsmtFinType1))
sum(is.na(house_sub$BsmtFinSF1))
sum(is.na(house_sub$BsmtFinType2))
sum(is.na(house_sub$BsmtFinSF2))
sum(is.na(house_sub$BsmtUnfSF))
sum(is.na(house_sub$TotalBsmtSF))
sum(is.na(house_sub$Heating))
sum(is.na(house_sub$SalePrice))
```


#Outliers

#Numerical Values
``` {r outliers}
sum(house_sub$YearRemodAdd < (quantile(house_sub$YearRemodAdd, .25) - 1.5*IQR(house_sub$YearRemodAdd)))
sum(house_sub$YearRemodAdd > (quantile(house_sub$YearRemodAdd, .75) + 1.5*IQR(house_sub$YearRemodAdd)))
sum(house_sub$BsmtFinSF1 < (quantile(house_sub$BsmtFinSF1, .25) - 1.5*IQR(house_sub$BsmtFinSF1)))
sum(house_sub$BsmtFinSF1 > (quantile(house_sub$BsmtFinSF1, .75) + 1.5*IQR(house_sub$BsmtFinSF1)))
sum(house_sub$BsmtFinSF2 < (quantile(house_sub$BsmtFinSF2, .25) - 1.5*IQR(house_sub$BsmtFinSF2)))
sum(house_sub$BsmtFinSF2 > (quantile(house_sub$BsmtFinSF2, .75) + 1.5*IQR(house_sub$BsmtFinSF2)))
sum(house_sub$BsmtUnfSF < (quantile(house_sub$BsmtUnfSF, .25) - 1.5*IQR(house_sub$BsmtUnfSF)))
sum(house_sub$BsmtUnfSF > (quantile(house_sub$BsmtUnfSF, .75) + 1.5*IQR(house_sub$BsmtUnfSF)))
sum(house_sub$TotalBsmtSF < (quantile(house_sub$TotalBsmtSF, .25) - 1.5*IQR(house_sub$TotalBsmtSF)))
sum(house_sub$TotalBsmtSF > (quantile(house_sub$TotalBsmtSF, .75) + 1.5*IQR(house_sub$TotalBsmtSF)))
sum(house_sub$SalePrice < (quantile(house_sub$SalePrice, .25) - 1.5*IQR(house_sub$SalePrice)))
sum(house_sub$SalePrice > (quantile(house_sub$SalePrice, .75) + 1.5*IQR(house_sub$SalePrice)))
```
#Categorical Values
``` {r na_sum}
apply(house_sub, MARGIN = 2, table)
```

