---
title: "EDA"
author: "chungmeng"
date: "October 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Initialization}
library(ggplot2)
library(dplyr)
library(tidyr)
setwd("C:/Users/cmlim/Dropbox/NYC Data Science/Class/Machine Learning Project/chungmeng")
house_train <- read.csv("../data/train.csv", 
                        header = TRUE, 
                        na.strings = "",
                        stringsAsFactors = FALSE)
```

##Columns of Interest

Extract Columns [41:60] & Price
```{r mycols}
myhouse_train=house_train[,41:60]
myhouse_train=cbind(myhouse_train,house_train[81])
names(myhouse_train)
```
## Missingness
```{r Missingness}

colSums(is.na(myhouse_train))
#colSums(myhouse_train=="")
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
myhouse_train$GarageYrBlt=as.numeric(myhouse_train$GarageYrBlt)
summary(myhouse_train)

```
## Square Foot Plots (4 Variables)
```{r}

ggplot(myhouse_train, aes(x=X1stFlrSF))+geom_histogram()+ggtitle("1st Floor Area")
ggplot(myhouse_train, aes(x=X2ndFlrSF))+geom_histogram()+ggtitle("2nd Floor Area")
ggplot(myhouse_train, aes(x=LowQualFinSF))+geom_histogram()+ggtitle("Low Ql Floor Area")
ggplot(myhouse_train, aes(x=GrLivArea))+geom_histogram()+ggtitle("Above Ground Living Area")

```
Useless: Low QL Floor Area. Nearly all 0's

## Room/Bath/Fireplace Counts (8 Variables)
```{r}
ggplot(myhouse_train, aes(x=BsmtFullBath))+geom_bar()
ggplot(myhouse_train, aes(x=BsmtHalfBath))+geom_bar()
ggplot(myhouse_train, aes(x=FullBath))+geom_bar()
ggplot(myhouse_train, aes(x=HalfBath))+geom_bar()
ggplot(myhouse_train, aes(x=BedroomAbvGr))+geom_bar()
ggplot(myhouse_train, aes(x=KitchenAbvGr))+geom_bar()
ggplot(myhouse_train, aes(x=TotRmsAbvGrd))+geom_bar()
ggplot(myhouse_train, aes(x=Fireplaces))+geom_bar()
```

## Quality Factors/Categorical (6 Variables)
```{r}
ggplot(myhouse_train, aes(x=HeatingQC))+geom_bar()
ggplot(myhouse_train, aes(x=Electrical))+geom_bar()
ggplot(myhouse_train, aes(x=CentralAir))+geom_bar()
ggplot(myhouse_train, aes(x=KitchenQual))+geom_bar()
ggplot(myhouse_train, aes(x=FireplaceQu))+geom_bar()
ggplot(myhouse_train, aes(x=GarageType))+geom_bar()
```
Electrical is just mostly SBrkr Type

#Year Built
```{r}
ggplot(myhouse_train, aes(x=GarageYrBlt))+geom_histogram()
```

##Numerical Columns
```{r mynumcols}
num_cols <- sapply(myhouse_train, is.numeric)
num_myhouse_train=myhouse_train[,num_cols]
```
## Scatter with SalePrice
```{r}
num_myhouse_train
plot(num_myhouse_train$X1stFlrSF,num_myhouse_train$SalePrice)
plot(num_myhouse_train$X2ndFlrSF,num_myhouse_train$SalePrice)
plot(num_myhouse_train$GrLivArea,num_myhouse_train$SalePrice)
```
Strong:
```{r}

ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$BsmtFullBath),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$BsmtHalfBath),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$FullBath),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$HalfBath),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$BedroomAbvGr),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(myhouse_train$KitchenAbvGr),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$TotRmsAbvGrd),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)
ggplot(num_myhouse_train, aes(x=as.factor(num_myhouse_train$Fireplaces),y=SalePrice))+geom_violin()+geom_boxplot(width=.1)


```
Kitchen Above Ground doesn't seem useful
Bedroom Above Ground doesn't have a clear relationship with Prices

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
