---
title: "Project 3 Missingness Analysis"
author: "Wing"
date: "10/31/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      warning = FALSE, 
                      message = FALSE,
                      tidy=FALSE,
                      fig.height=6,
                      fig.width=10)

```

```{r initialization}
library(ggplot2)
library(dplyr)
library(data.table)
library(VIM)
library(caret)
library(RANN)
```
## DATA MUNGING

1. Introduction
2. Loading Data
3. Visualizations
4. Pre-Processing
5. Model Training and Parameter Tuning
6. Variable Importance and Feature selection
7. Summary


###  Loading data from csv

```{r load_data}
## load training data
house_train <- read.csv("./train.csv", 
                        header = TRUE, 
                        na.strings = c("", "NA"),
                        stringsAsFactors = FALSE)
## load test data
house_test <- read.csv("./test.csv", 
                        header = TRUE, 
                        na.strings = c("", "NA"),
                        stringsAsFactors = FALSE)
dim(house_train)
dim(house_test)
str(house_train)
```

### Getting factor levels from data description

```{r factor_levels}
## get levels of categorical features from data description
factorLevel <- list()
conn <- file("data_description.txt", open="r")
f <-readLines(conn)
for (line in f){
  if(!grepl("^[[:blank:]]", line) & grepl(": ", line)) {
    col_name <<- trimws(gsub(":.*", "", line))
  } else {
    level <- trimws(gsub("\t.*", "", line))
    if (level != "") {
      factorLevel[[col_name]] <- c(factorLevel[[col_name]], level)
    }
  }
}
close(conn)

print(factorLevel[1:6])
```

###Checking factor levels with data

```{r check_levels}
## check if levels in description cover unique data values
for (varname in names(factorLevel)) {
  levelDiff <- setdiff(unique(house_train[[varname]]), 
                       factorLevel[[varname]])
  if(length(levelDiff)) {
    print(paste(varname, 
                paste(levelDiff, collapse = ", "), 
                sep = ": "))
  }
}

```


### Fixing level names

```{r fix_levels}
## fix those levels that don't match with data
## ignore "NA" as they will be considered as missing when converting categorical to factors

unique(house_train$MSZoning)
factorLevel$MSZoning
factorLevel$MSZoning[2] <- "C (all)"

unique(house_train$Neighborhood)
factorLevel$Neighborhood
factorLevel$Neighborhood[13] <- "NAmes"

unique(house_train$BldgType)
factorLevel$BldgType
factorLevel$BldgType[c(2,3,5)] <- c("2fmCon","Duplex","Twnhs")

unique(house_train$Exterior2nd)
factorLevel$Exterior2nd
factorLevel$Exterior2nd[c(17,6,3)] <- c("Wd Shng","CmentBd","Brk Cmn")

## Get levels that only appear in the dataset
for (varname in names(factorLevel)) {
  factorLevel[[varname]] <- intersect(factorLevel[[varname]],
                                      unique(house_train[[varname]]))
}

## Re-run the previous cell to double check
```

### Converting column data types

```{r convert_types}
## convert column datatype to numeric / factor
## On training dataset
for (varname in names(house_train)[-1]) {
  if (varname %in% names(factorLevel)) {
    house_train[[varname]] <- factor(house_train[[varname]], 
                                     levels = factorLevel[[varname]])
  } else {
    house_train[[varname]] <- as.numeric(house_train[[varname]])
  }
}

## On testing dataset
for (varname in names(house_test)[-1]) {
  if (varname %in% names(factorLevel)) {
    house_test[[varname]] <- factor(house_test[[varname]], 
                                    levels = factorLevel[[varname]])
  } else {
    house_test[[varname]] <- as.numeric(house_test[[varname]])
  }
}
```


### SAVING DATA: These two data frames (test and train) contain all variables.

```{r save_loaded}
house_train$Id <- NULL
rownames(house_test) <- house_test$Id
house_test$Id <- NULL
save(house_test, file = "house_test.Rda" )
save(house_train, file =  "house_train.Rda")
```

##EXAMINE MISSING VALUES

###Create subset of training set to include a smaller list of variables decided on 11/1.

```{r create_subset}

name_vec <- c("MSZoning", "Neighborhood", "HouseStyle", "OverallQual", "YearBuilt", "ExterQual",
              "Foundation", "BsmtQual", "X1stFlrSF", "X2ndFlrSF", "GrLivArea", "FullBath",
              "TotRmsAbvGrd" ,"GarageFinish", "GarageCars", "GarageArea", "GarageQual",
              "GarageCond", "PavedDrive", "SaleType", "SaleCondition")


train_sub <- house_train %>% select(., name_vec)
```

###Count missing values
####Bsmt and the Garage variables are the only ones with missing values.
```{r check_missingness}

col_missing <- names(train_sub[colSums(is.na(train_sub)) > 0])
missing_df <- train_sub[,col_missing]
aggr(missing_df)
Filter(function(x) x > 0, colSums(is.na(missing_df)))
```


####Inspect missing observations. 
It appears that all the garage variables have the same observations with missing value.

```{r inspect_missing}
missing_df <- train_sub[rowSums(is.na(train_sub))>0,] #checking all observations with missing
train_sub[rowSums(is.na(train_sub))>3,] #checking observations with all 4 variables missing
table()
```

### Relabel missing values with "None"
```{r relabel}
train_sub <- data.table(train_sub)
train_sub[is.na(train_sub$BsmtQual), BsmtQual := "None"] 
train_sub[is.na(train_sub$GarageFinish), GarageFinish := "None"] 
train_sub[is.na(train_sub$GarageQual), GarageQual := "None"] 
train_sub[is.na(train_sub$GarageCond), GarageCond := "None"]
train_sub[PavedDrive == "N", PavedDrive := "None"] 

#check levels
levels_sub = sapply(train_sub, function (x) levels(x))
```

### SAVING DATA
```{r save_train_sub}
save(train_sub, file = "train_sub.Rda")
```

### Inspect Redundant Variables
GarageCars captures all the missing values in the other Garage variables as well.
```{r redundant_variables}

x <-  train_sub[GarageFinish == "None" & GarageCars== 0,]
table(train_sub$GarageCars, train_sub$GarageFinish)

```